---
title: "Weather Event Impacts on Public Health & The Economy"
author: "Sam Edwardes"
date: "February 18, 2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Synopsis
The purpose of this document is to address the following questions:

1. Across the United States, which types of weather events are most harmful with respect to population health?
2. Across the United States, which types of weather events have the greatest economic consequences?

Based on the analysis, it has been determined that Tornados have the biggest impact on population health, and floods have the biggest economic impact.


## Data Processing

### Load the data
The data can be downloaded from the URL below and then loaded into R. 

[Weather data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)
```{r message=FALSE}
library(data.table)
library(dplyr)
library(tidyr)
library(R.utils)
library(ggplot2)
library(gridExtra)
```
```{r cache=TRUE}
# Data can be downloaded from the following URL

url1 <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
file_name1 <- "repdata_data_StormData.csv.bz2"
file_name2 <- "repdata_data_StormData.csv"


# Download and unzip the file

if(!dir.exists(file_name1)){download.file(url = url1, destfile = file_name1)}
if(!dir.exists(file_name2)){bunzip2(file_name1, "repdata_data_StormData.csv", remove = FALSE, skip = TRUE)}


# Load the data

df <- fread(file_name2)
```
### Review the data
```{r}
# Examine the data in EVTYPE and the resulting health outcomes

df %>%
    group_by(EVTYPE) %>%
    summarise(
        NUM_RECORDS = n(),
        FATALITIES = sum(FATALITIES),
        INJURIES = sum(INJURIES),
        PROPDMG = sum(PROPDMG),
        CROPDMG = sum(CROPDMG)
    )
```


### Clean and subset the data
Clean the data.
```{r}
# Select required columns only, remove columns with no data

df_clean <- df %>%
    select(EVTYPE, FATALITIES, INJURIES, PROPDMG, PROPDMGEXP, CROPDMG, CROPDMGEXP) %>%
    filter(FATALITIES > 0 | INJURIES > 0 | PROPDMG > 0 | CROPDMG > 0) %>%
    mutate(
        PROPDMGEXP = toupper(PROPDMGEXP),
        CROPDMGEXP = toupper(CROPDMGEXP),
        PROPDMG_CLEAN = case_when(
            PROPDMGEXP == 2 ~ PROPDMG * 10,
            PROPDMGEXP == 3 ~ PROPDMG * 100,
            PROPDMGEXP == 4 ~ PROPDMG * 1000,
            PROPDMGEXP == 5 ~ PROPDMG * 10000,
            PROPDMGEXP == 6 ~ PROPDMG * 100000,
            PROPDMGEXP == 7 ~ PROPDMG * 1000000,
            PROPDMGEXP == 8 ~ PROPDMG * 10000000,
            PROPDMGEXP == 9 ~ PROPDMG * 100000000,
            PROPDMGEXP == "H" ~ PROPDMG * 10,
            PROPDMGEXP == "K" ~ PROPDMG * 100,
            PROPDMGEXP == "M" ~ PROPDMG * 100000,
            PROPDMGEXP == "B" ~ PROPDMG * 1000000000,
            TRUE ~ PROPDMG
        ),
        CROPDMG_CLEAN = case_when(
            CROPDMGEXP == "K" ~ CROPDMG * 100,
            CROPDMGEXP == "M" ~ CROPDMG * 100000,
            CROPDMGEXP == "B" ~ CROPDMG * 1000000000,
            TRUE ~ CROPDMG
        )
    )
```
## Results
***Q1: Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?***

The data contains two health outcomes:

1. Fatalities
2. Injuries

While both are bad, in order to understand what is most harmful to population health, we can assign a weighting to each. This analysis will assume that a fatality is twice as bad as an injury. This logic will be used to create a field `HEALTH SCORE`. The higher the score, the worse the impact is on population health.
```{r}
# Create the HEALTH_SCORE field

df_clean <- df_clean %>% mutate(HEALTH_SCORE = INJURIES + FATALITIES*2 )


# Examine the data in EVTYPE and the resulting health outcomes

df_clean_health <- df_clean %>%
    group_by(EVTYPE) %>%
    summarise(
        NUM_RECORDS = n(),
        FATALITIES = sum(FATALITIES),
        INJURIES = sum(INJURIES),
        HEALTH_SCORE = sum(HEALTH_SCORE)) %>%
    mutate(HEALTH_SCORE_PERCENT = (HEALTH_SCORE/sum(df_clean$HEALTH_SCORE))) %>%
    arrange(desc(HEALTH_SCORE)) %>%
    mutate(HEALTH_SCORE_PERCENT_CUM = cumsum(HEALTH_SCORE_PERCENT))

head(df_clean_health,20)


# Plot the results

df_clean_health_plot <- head(df_clean_health,10) %>%
    select(EVTYPE, FATALITIES, INJURIES, TOTAL_HEALTH_SCORE = HEALTH_SCORE) %>%
    gather(key = "key", value = "Value", -EVTYPE)

health_plot <- ggplot(df_clean_health_plot, aes(x=reorder(EVTYPE, -Value), y=Value)) +
    geom_bar(stat = "identity", aes(fill=key), position="dodge") + 
    theme(axis.text.x = element_text(angle=45, hjust=1)) + 
    labs(title="Top 10 US Weather Events by Health Score") +
    xlab("Weather Event")

health_plot
```

As the table shows us, Tornados are by far the biggest risk to public health, responsible for over 60% of the damage to population health. Additionally, the top 10 events make up more than 87% of all health impact.

**Q2: Across the United States, which types of events have the greatest economic consequences?**
```{r}
# Examine the data in EVTYPE and the resulting economic outcomes

df_clean_econ <- df_clean %>%
    group_by(EVTYPE) %>%
    mutate(TOTALDMG = PROPDMG_CLEAN + CROPDMG_CLEAN) %>%
    summarise(
        NUM_RECORDS = n(),
        PROPDMG_CLEAN = sum(PROPDMG_CLEAN),
        CROPDMG_CLEAN = sum(CROPDMG_CLEAN),
        TOTALDMG = sum(TOTALDMG)) %>%
    mutate(TOTALDMG_PERCENT = (TOTALDMG/(sum(df_clean$PROPDMG_CLEAN)+sum(df_clean$CROPDMG_CLEAN)))) %>%
    arrange(desc(TOTALDMG)) %>%
    mutate(TOTALDMG_CUM = cumsum(TOTALDMG_PERCENT))

head(df_clean_econ,20)


# Plot the results

df_clean_econ_plot <- head(df_clean_econ,10) %>%
    select(EVTYPE, PROPDMG_CLEAN, CROPDMG_CLEAN, TOTALDMG) %>%
    gather(key = "key", value = "Value", -EVTYPE)

econ_plot <- ggplot(df_clean_econ_plot, aes(x=reorder(EVTYPE, -Value), y=Value)) +
    geom_bar(stat = "identity", aes(fill=key), position="dodge") + 
    theme(axis.text.x = element_text(angle=45, hjust=1)) + 
    labs(title="Top 10 US Weather Events by Economic Impact") +
    xlab("Weather Event")

econ_plot
```

Floods have by far the greatest economic impact, accounting for more than 40% of damage. The top 10 events account for more than 91% of all economic damage.

### Conclusion
Based on the analysis, it has been determined that Tornados have the biggest impact on population health, and floods have the biggest economic impact.
```{r fig.width=14}
grid.arrange(health_plot, econ_plot, ncol = 2)
```