---
title: 'Assignment Part 2: Basic Inferential Data Analysis'
author: "Sam Edwardes"
date: "April 5, 2019"
output:
  html_document:
    df_print: paged
---
### Overview
The purpose of this document is to analyze tooth growth data in the R datasets package. From the R Documentation (?ToothGrowth):

*The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC).*

*The dataframe has 60 observations on 3 variables:*

- *[,1]	 len	 numeric	 Tooth length*
- *[,2]	 supp	 factor	 Supplement type (VC or OJ)*
- *[,3]	 dose	 numeric	 Dose in milligrams/day*

### Load data
```{r message=FALSE}
# load required libraries
library(ggplot2)
library(dplyr)
library(Hmisc)

# load ToothGrowth data into a dataframe
tg <- datasets::ToothGrowth
```

### Exploratory Analysis
First, lets explore what the data looks like, and run some basic summary statistics.
```{r}
describe(tg)
```

From analyzing each column, there are several insights gleaned:

- [len] column has the most variability, with values range from 4.2 up to 33.9
- [supp] column has only two variables, OJ and VC
- [does] column has three potential values, either 0.5, 1, or 2

It looks like we will want to compare tooth growth based on delivery method and/or dose. Lets first compare tooth growth by delivery method.
```{r}
g1 <- ggplot(tg, aes(y = len, x = supp)) +
    geom_boxplot(outlier.shape = NA, width = 0.3) +
    stat_summary(aes(ymax = ..y.., ymin = ..y..),fun.y = mean, color='blue', geom="errorbar", linetype = "dashed") + 
    geom_point(alpha = 0.2, size = 3) +
    labs(title = "Tooth growth comparing delivery methods")

g1
```

From the plot above, it looks like orange juice (OJ) results in higher tooth growth. Lets compare tooth growth by dose.
```{r}
g2 <- ggplot(tg, aes(y = len, x = as.character(dose))) +
    geom_boxplot(outlier.shape = NA, width = 0.3) +
    stat_summary(aes(ymax = ..y.., ymin = ..y..),fun.y = mean, color='blue', geom="errorbar", linetype = "dashed") + 
    geom_point(alpha = 0.2, size = 3) +
    labs(title = "Tooth growth comparing delivery methods", x = "dose")

g2
```

As you would expect, as the dose increases, so does tooth growth. Lets compare toothgrowth by dose and delivery method.
```{r}
g3 <- ggplot(tg, aes(y = len, x = as.character(dose))) +
    geom_boxplot(outlier.shape = NA, width = 0.3) +
    stat_summary(aes(ymax = ..y.., ymin = ..y..),fun.y = mean, color='blue', geom="errorbar", linetype = "dashed") + 
    geom_point(alpha = 0.2, size = 3) +
    labs(title = "Tooth growth comparing dose and delivery methods", x = "dose") +
    facet_grid(.~supp)

g3
```

As the chart above demonstrates:

- it looks like at the highest dose, both delivery methods result in similar results
- at lower doeses (0.5 and 1.0), orange juice delivers superior results.

### Confidence internvals
Now lets confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose to see if our initial observations are valid.

### Conclusions
...