---
title: "Quiz 2"
author: "Sam Edwardes"
date: "March 24, 2019"
output: html_document
---

#1
What is the variance of the distribution of the average an IID draw of nnn observations from a population with mean mu\muμ and variance sigma, start superscript, 2, end superscript\sigma^2σ

#2
Suppose that diastolic blood pressures (DBPs) for men aged 35-44 are normally distributed with a mean of 80 (mm Hg) and a standard deviation of 10. About what is the probability that a random 35-44 year old has a DBP less than 70?
```{r}
mu <- 80
sigma <- 10
q <- 70

pnorm(q, mean = mu, sd = sigma, lower.tail = TRUE)
```

#3
Brain volume for adult women is normally distributed with a mean of about 1,100 cc for women with a standard deviation of 75 cc. What brain volume represents the 95th percentile?
```{r}
mu <- 1100
sigma <- 75
p <- 0.95

qnorm(p, mu, sigma)
```
#4
Refer to the previous question. Brain volume for adult women is about 1,100 cc for women with a standard deviation of 75 cc. Consider the sample mean of 100 random adult women from this population. What is the 95th percentile of the distribution of that sample mean?
```{r}
1100 + c(-1,1) *qnorm(0.95) * 75/sqrt(100)
```

#5
You flip a fair coin 5 times, about what's the probability of getting 4 or 5 heads?
```{r}
# probablyility of 4 heads
n <- 5
p <- 0.5

ways_to_win <- choose(5,4)
outcomes <- 2^5
heads_4 <- ways_to_win / outcomes
heads_4

# probability of 5 heads
ways_to_win <- 1
heads_5 <- ways_to_win / outcomes
heads_5

# probabilyt of 4 or 5 heads
heads_4 + heads_5
```
#6
The respiratory disturbance index (RDI), a measure of sleep disturbance, for a specific population has a mean of 15 (sleep events per hour) and a standard deviation of 10. They are not normally distributed. Give your best estimate of the probability that a sample mean RDI of 100 people is between 14 and 16 events per hour?
```{r}
pnorm(16, mean = 15, sd = 1) - pnorm(14, mean = 15, sd = 1)
```
#7
Consider a standard uniform density. The mean for this density is .5 and the variance is 1 / 12. You sample 1,000 observations from this distribution and take the sample mean, what value would you expect it to be near?
```{r}
mu <- 0.5
var <- 1/12
sigma <- (1/12)^2
n <- 1000

```
#8
The number of people showing up at a bus stop is assumed to be

Poisson with a mean of 5 5 5 people per hour. You watch the bus

stop for 3 hours. About what's the probability of viewing 10 or fewer people?
```{r}
mu <- 5
t <- 3

ppois(10, mu*t )
```

