# Quiz Week 2

## Q1
Load the required packages for the quiz
```{r}
library(AppliedPredictiveModeling)
data(segmentationOriginal)
library(caret)
```
Load the libs I need
```{r}
library(rattle)
library(rpart.plot)
```

Steps to load the data
### Step 1
Subset the data to a training set and testing set based on the Case variable in the data set.
```{r}
testing <- segmentationOriginal[segmentationOriginal$Case == "Test",]
training <- segmentationOriginal[segmentationOriginal$Case == "Train",]
```

### Step 2
Set the seed to 125 and fit a CART model to predict Class with the rpart method using all predictor variables and default caret settings.
```{r}
set.seed(125)

# fit CART model
fit <- train(Class ~ ., data = training, method = "rpart")
```


### Step 3
In the final model what would be the final model prediction for cases with the following variable values:

```{r}
newdata_a <- data.frame(TotalIntench2 = 23000,
                        FiberWidthCh1 = 10, 
                        PerimStatusCh1 = 2)

# predict(fit, newdata = newdata_a, type = "prob")
fancyRpartPlot(fit$finalModel)
```

### Answers

a. TotalIntench2 = 23,000; FiberWidthCh1 = 10; PerimStatusCh1=2

**PS**

b. TotalIntench2 = 50,000; FiberWidthCh1 = 10;VarIntenCh4 = 100

**WS**

c. TotalIntench2 = 57,000; FiberWidthCh1 = 8;VarIntenCh4 = 100

**PS**

d. FiberWidthCh1 = 8;VarIntenCh4 = 100; PerimStatusCh1=2

**not possible to predict**

## Q2

If K is small in a K-fold cross validation is the bias in the estimate of out-of-sample (test set) accuracy smaller or bigger? *larger bias*

If K is small is the variance in the estimate of out-of-sample (test set) accuracy smaller or bigger. *smaller*

Is K large or small in leave one out cross validation? *small*

- The bias is larger and the variance is smaller. Under leave one out cross validation K is equal to two.
- The bias is smaller and the variance is bigger. Under leave one out cross validation K is equal to one.
- The bias is smaller and the variance is smaller. Under leave one out cross validation K is equal to one.
- **The bias is larger and the variance is smaller. Under leave one out cross validation K is equal to the sample size.**

## Q3
```{r}
library("pgmm")
data(olive)
olive = olive[,-1]
```

These data contain information on 572 different Italian olive oils from multiple regions in Italy. Fit a classification tree where Area is the outcome variable. Then predict the value of area for the following data frame using the tree command with all defaults

```{r}
# model fil
fit <- train(Area ~ .,
             data = olive,
             method = "rpart")

# test data
newdata = as.data.frame(t(colMeans(olive)))

# predict
predict(fit, newdata = newdata)

# answer = 2.783
```

## Q4

load the data
```{r}
library("ElemStatLearn")
data(SAheart)
set.seed(8484)
train = sample(1:dim(SAheart)[1],size=dim(SAheart)[1]/2,replace=F)
trainSA = SAheart[train,]
testSA = SAheart[-train,]

# from courera
missClass = function(values,prediction){sum(((prediction > 0.5)*1) != values)/length(values)}
```

Then set the seed to 13234 and fit a logistic regression model (method="glm", be sure to specify family="binomial") with Coronary Heart Disease (chd) as the outcome and age at onset, current alcohol consumption, obesity levels, cumulative tabacco, type-A behavior, and low density lipoprotein cholesterol as predictors. Calculate the misclassification rate for your model using this function and a prediction on the "response" scale:

```{r}
# set seed
set.seed(13234)

fit <- train(chd ~ age + alcohol + obesity + tobacco + typea + ldl,
             data = trainSA,
             method = "glm",
             family = "binomial")


# answer
missClass(testSA$chd, predict(fit, newdata = testSA)) # 0.281
missClass(trainSA$chd, predict(fit, newdata = trainSA)) # 0.311
```

## Q4 v2

Load the South Africa Heart Disease Data and create training and test sets with the following code:

```{r}
library(ElemStatLearn)
data(SAheart)
set.seed(8484)
train = sample(1:dim(SAheart)[1],size=dim(SAheart)[1]/2,replace=F)
trainSA = SAheart[train,]
testSA = SAheart[-train,]
```

Then set the seed to 13234 and fit a logistic regression model (method="glm", be sure to specify family="binomial") with Coronary Heart Disease (chd) as the outcome and age at onset, current alcohol consumption, obesity levels, cumulative tabacco, type-A behavior, and low density lipoprotein cholesterol as predictors. Calculate the misclassification rate for your model using this function and a prediction on the "response" scale:

```{r}
missClass = function(values,prediction){sum(((prediction > 0.5)*1) != values)/length(values)}
```

What is the misclassification rate on the training set? What is the misclassification rate on the test set?
### Answer

```{r, fig.height=15}
set.seed(13234)
fit <- train(chd ~ age + alcohol + obesity + tobacco + typea + ldl, data=trainSA, method="glm", family="binomial")
predictTrainSA <- predict(fit)
missClass(trainSA$chd,predictTrainSA)
predictTestSA <- predict(fit, testSA)
missClass(testSA$chd,predictTestSA)
```

## Q4 v3
```{r}
library(ElemStatLearn)
data(SAheart)
set.seed(8484)
train = sample(1:dim(SAheart)[1], size = dim(SAheart)[1] / 2, replace = F)
trainSA = SAheart[train, ]
testSA = SAheart[-train, ]

missClass = function(values, prediction){sum(((prediction > 0.5) * 1) != values) / length(values)}

set.seed(13234)
modelSA <- train(chd ~ age + alcohol + obesity + tobacco + typea + ldl, 
               data = trainSA, method = "glm", family = "binomial")

missClass(testSA$chd, predict(modelSA, newdata = testSA))
missClass(trainSA$chd, predict(modelSA, newdata = trainSA))
```

## Q5
```{r}
library(ElemStatLearn)
data(vowel.train)
data(vowel.test)
```

Set the variable y to be a factor variable in both the training and test set. Then set the seed to 33833. Fit a random forest predictor relating the factor variable y to the remaining variables. Read about variable importance in random forests here: http://www.stat.berkeley.edu/~breiman/RandomForests/cc_home.htm#ooberr The caret package uses by default the Gini importance.

Calculate the variable importance using the varImp function in the caret package. What is the order of variable importance?

[NOTE: Use randomForest() specifically, not caret, as there's been some issues reported with that approach. 11/6/2016]

```{r}
library("randomForest")

# set Y to a factor
vowel.test$y <- factor(vowel.test$y)
vowel.train$y <- factor(vowel.train$y)

set.seed(33833)

# fit the model
fit <- randomForest(y ~ .,
                    data = vowel.train)

order(varImp(fit), decreasing = TRUE)

```
Answer: 
MeanDecreaseGini
x.1          87.27808
x.2          87.26216
x.3          33.71495
x.4          36.71464
x.5          51.98077
x.6          45.38691
x.7          31.15063
x.8          42.12067
x.9          33.64751
x.10         29.83885

 x.2, x.1, x.5, x.6, x.8, x.4, x.9, x.3, x.7, x.10
