# Quiz Week 2

## Q #1
```{r}
library(AppliedPredictiveModeling)
data(AlzheimerDisease)
```
## Q #2
```{r fig.height=10, fig.width=10}
library(AppliedPredictiveModeling)
data(concrete)
library(caret)
set.seed(1000)
inTrain = createDataPartition(mixtures$CompressiveStrength, p = 3/4)[[1]]
training = mixtures[ inTrain,]
testing = mixtures[-inTrain,]

# Make a plot of the outcome (CompressiveStrength) versus the index of the samples. Color by each of the variables in the data set (you may find the cut2() function in the Hmisc package useful for turning continuous covariates into factors). What do you notice in these plots?
library(ggplot2)
library(GGally)
library(Hmisc)

training_raw <- training

training$index <- rownames(training)
training$CompressiveStrength_bin <- cut2(training$CompressiveStrength, g = 4)
training$Age_bin <- cut2(training$Age, g = 4)
training$FlyAsh_bin <- cut2(training$FlyAsh, g = 4)
# 
# ggpairs(data = training,
#         columns = c("FlyAsh","Age","CompressiveStrength"),
#         mapping = aes(colour = CompressiveStrength_bin))
# 
# # pairs plot
# featurePlot(x=training, y=training$CompressiveStrength_bin, plot = "pairs")

# outcome, index, age
g1 <- ggplot(data = training, aes(x = index, y = CompressiveStrength, colour = Age_bin)) +
    geom_point() + 
    ggtitle("#1 Compressive Strength vs. Index/Age")
g1 # does not explain pattern

# outome vs. index
# outcome, index, age
g3 <- ggplot(data = training, aes(x = index, y = CompressiveStrength)) +
    geom_point() + 
    ggtitle("#3 Compressive Strength vs. Index")
g3 # does not explain pattern

# outcome, index, age
g4 <- ggplot(data = training, aes(x = index, y = CompressiveStrength, colour = FlyAsh_bin)) +
    geom_point() + 
    ggtitle("#4 Compressive Strength vs. Index/Age")
g4 # does not explain pattern

## ggpairs
training2 <- training_raw
training2$CompressiveStrength_bin <- cut2(training$CompressiveStrength, g = 3)
ggpairs(data = training2, mapping = ggplot2::aes(colour = CompressiveStrength_bin), progress = FALSE, axisLabels = "internal")

# There is a non-random pattern in the plot of the outcome versus index that does not appear to be perfectly explained by any predictor suggesting a variable may be missing.

```
## Q #3
```{r}
# load the data
library(AppliedPredictiveModeling)
data(concrete)
library(caret)
set.seed(1000)
inTrain = createDataPartition(mixtures$CompressiveStrength, p = 3/4)[[1]]
training = mixtures[ inTrain,]
testing = mixtures[-inTrain,]

# clean data
training2 <- training

# Make a histogram and confirm the SuperPlasticizer variable is skewed. Normally you might use the log transform to try to make the data more symmetric. Why would that be a poor choice for this variable?
g1 <- ggplot(data = training2) +
    geom_histogram(aes(Superplasticizer))
g1

# using log
training3 <- training
training3$Superplasticizer <- log(training3$Superplasticizer)
g2 <- ggplot(data = training3) +
    geom_histogram(aes(Superplasticizer))
g2

# bad to log because big tail at 0
```

## Q #4
```{r}
library(caret)
library(AppliedPredictiveModeling)
set.seed(3433)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]

# Find all the predictor variables in the training set that begin with IL. Perform principal components on these variables with the preProcess() function from the caret package. Calculate the number of principal components needed to capture 90% of the variance. How many are there?

pattern_match <- grepl(pattern = "^IL", x = names(testing))
trainingIL <- training[,pattern_match]
# training2 <- bind_cols(training2, as.data.frame(training$diagnosis))
str(training2)

print("########## Pre Process ############")
# perform PCA
prePro <- preProcess(x = trainingIL,
                  method = "pca",
                  thresh = 0.9)
prePro

# 9 components required

```

## Q #5
Create a training data set consisting of only the predictors with variable names beginning with IL and the diagnosis. Build two predictive models, one using the predictors as they are and one using PCA with principal components explaining 80% of the variance in the predictors. Use method="glm" in the train function.

What is the accuracy of each method in the test set? Which is more accurate?
```{r}
library(caret)
library(AppliedPredictiveModeling)
set.seed(3433)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]

# keep only variables names beginning with IL and diagnosis
pattern_match <- grepl(pattern = "^IL|diagnosis", x = names(testing))
training_IL <- training[,pattern_match]
testing_IL <- testing[,pattern_match]


# predictive model 1, non-PCA
model <- train(diagnosis ~ .,
               data = training_IL,
               method = "glm")
predict_model <- predict(model, newdata = testing_IL)
matrix_model <- confusionMatrix(predict_model, testing_IL$diagnosis)
matrix_model$overall[1]


# predictive model 2, with PCA
modelPCA <- train(diagnosis ~ .,
                  data = training_IL,
                  preProcess = "pca",
                  trControl=trainControl(preProcOptions=list(thresh=0.8)))
predict_model_PCA <- predict(modelPCA, newdata = testing_IL)
matrix_model_PCA <- confusionMatrix(predict_model_PCA, testing_IL$diagnosis)
matrix_model_PCA$overall[1]

#65 and 72

```
## Q #5 attempt two
```{r}
library(caret)
library(AppliedPredictiveModeling)
set.seed(3433)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[ inTrain,]
testing = adData[-inTrain,]

# grep all columns with IL and diagnosis in the traning and testing set
trainingIL <- training[,grep("^IL|diagnosis", names(training))]
testingIL <- testing[,grep("^IL|diagnosis", names(testing))]

# non-PCA
model <- train(diagnosis ~ ., data = trainingIL, method = "glm")
predict_model <- predict(model, newdata= testingIL)
matrix_model <- confusionMatrix(predict_model, testingIL$diagnosis)
matrix_model$overall[1]

# PCA
modelPCA <- train(diagnosis ~., data = trainingIL, method = "glm", preProcess = "pca",trControl=trainControl(preProcOptions=list(thresh=0.8)))
matrix_modelPCA <- confusionMatrix(testingIL$diagnosis, predict(modelPCA, testingIL))
matrix_modelPCA$overall[1]
```


# Other